### 2014-12-03 writing/F'14 paper/discussion.tex
Tentative results in paper

This paper first evaluates two recursive models on a series of three increasingly
challenging interpretive tasks involving natural language inference---the 
core relational algebra of natural logic with entailment and
exclusion; recursive propositional logic structures; and statements
involving quantification and negation. We then showed that the same models can learn to
perform an entailment task on natural language image captions. \mynote{The results suggest that RNTNs,
but not necessarily plain RNNs}, have the capacity to model these tasks with 
reasonably-sized training sets. These positive results are
promising for the future of learned representation models in the
applied modeling of compositional semantics.

Of course, challenges remain. In terms of our experimental data, even
the RNTN falls short of perfection in the recursion experiment, with
performance falling off steadily as the size of the expressions grows. It
remains to be seen whether these deficiencies can be overcome with
stronger models or optimization procedures. The space of possible stronger
models is sizable, including extensions of the recursive models used here
\cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep}, recurrent
models \cite{sutskever2014sequence}, and even learning-enabled variants 
of more structured models like those of \newcite{grefenstette2013towards} or \newcite{rocktaschellow}.
In addition,
there remain subtle questions about how to fairly assess whether these
models have truly generalized in the way we want them to. There is a
constant tension between showing the models training data that gives
them a chance to learn the target logical functions and revealing the
answer to them in a way that leads to overfitting. The underlying
logical theories provide only limited guidance on this point.

>>>>

This paper first evaluates two recursive models on a series of three increasingly
challenging interpretive tasks involving natural language inference---the 
core relational algebra of natural logic with entailment and
exclusion; recursive propositional logic structures; and statements
involving quantification and negation. We then showed that the same models can learn to
perform an entailment task on natural language image captions. \mynote{The results suggest that RNTNs,
and potentially also RNNs}, have the capacity to model these tasks with 
reasonably-sized training sets. These positive results are
promising for the future of learned representation models in the
applied modeling of compositional semantics.

Of course, challenges remain. Even
the RNTN falls short of perfection in the recursion experiment, with
performance falling off steadily as the size of the expressions grows. It
remains to be seen whether these deficiencies can be overcome with
stronger models or optimization procedures. The space of possible stronger
models is sizable, including extensions of the recursive models used here
\cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep}, recurrent
models \cite{sutskever2014sequence}, and even learning-enabled variants 
of more structured models like those of \newcite{grefenstette2013towards} or \newcite{rocktaschellow}.
In addition,
there remain subtle questions about how to fairly assess whether these
models have truly generalized in the way we want them to. There is a
constant tension between showing the models training data that gives
them a chance to learn the target logical functions and revealing the
answer to them in a way that leads to overfitting. The underlying
logical theories provide only limited guidance on this point.

### 2014-12-03 writing/F'14 paper/intro.tex
Tentative results in paper

In our first set of experiments, we use this pre-specified logical
grammar to generate controlled data sets encoding the semantic
relationships between pairs of expressions and evaluate whether each
of two classes of neural model---plain recursive NNs and recursive neural
tensor networks (RNTNs; \citealt{socher2013acl1})---can learn those
relationships correctly. In our first experiment
(Section~\ref{sec:join}), we evaluate the ability of these models to
learn the core relational algebra of natural logic from data. Our
second experiment (Section~\ref{sec:recursion}) extends this
evaluation to relations between complex recursive structures like
$(a \plor b)$ and $\plneg (\plneg a \pland \plneg b)$, and our third
experiment (Section~\ref{sec:quantifiers}) involves relations between
quantified statements like \ii{every reptile walks} and \ii{no turtle
  moves}. We find that the plain recursive NN achieves only
  mixed results in all three experiments, whereas the stronger RNTN
  model generalizes well in every case, suggesting that it has in fact
  learned, or at least learned to simulate, our target logical
  concepts.

The experiments with simulated data provide a convincing demonstration
of the ability of neural networks to learn to encode semantic
inferences in complex natural language sentences from reasonably-sized
training sets. However, we are also interested in the more practical
question of whether they can learn these representations from more
naturalistic text. To address this question, we apply our models to
the SICK entailment challenge data. The small size of this corpus puts
data-intensive neural models like ours at a disadvantage, but we are
nonetheless able to achieve competitive performance on it
\mynote{CONSIDER ADDING A CONCRETE DETAIL --- E.G., ACCURACY
  AND ITS RELATION TO THE COMPETITION MEDIAN}, suggesting that the
representational abilities that we observe in the previous sections
are not limited to carefully circumscribed tasks. We conclude that RNTN models
are adequate for typical cases of natural logic inference, and that there is not yet
any clear level of inferential complexity for which other approaches work and NN models fail.

>>>>

In our first set of experiments, we use this pre-specified logical
grammar to generate controlled data sets encoding the semantic
relationships between pairs of expressions and evaluate whether each
of two classes of neural model---plain recursive NNs and recursive neural
tensor networks (RNTNs; \citealt{socher2013acl1})---can learn those
relationships correctly. In our first experiment
(Section~\ref{sec:join}), we evaluate the ability of these models to
learn the core relational algebra of natural logic from data. Our
second experiment (Section~\ref{sec:recursion}) extends this
evaluation to relations between complex recursive structures like
$(a \plor b)$ and $\plneg (\plneg a \pland \plneg b)$, and our third
experiment (Section~\ref{sec:quantifiers}) involves relations between
quantified statements like \ii{every reptile walks} and \ii{no turtle
  moves}. \mynote{Though the performance of the plain RNN model 
 is somewhat poor in one experiment, we find that the stronger RNTN
  model generalizes well in every case, suggesting that it has in fact
  learned, or at least learned to simulate, our target logical
  concepts.}

The experiments with simulated data provide a convincing demonstration
of the ability of neural networks to learn to encode semantic
inferences in complex natural language sentences from reasonably-sized
training sets. However, we are also interested in the more practical
question of whether they can learn these representations from more
naturalistic text. To address this question, we apply our models to
the SICK entailment challenge data. The small size of this corpus puts
data-intensive neural models like ours at a disadvantage, but we are
nonetheless able to achieve competitive performance on it,
\mynote{surpassing several models with significant hand engineered task specific features.}
This suggests that the representational abilities that we observe in the previous sections
are not limited to carefully circumscribed tasks. We conclude that RNTN models
are adequate for typical cases of natural logic inference, and that there is not yet
any clear level of inferential complexity for which other approaches work and NN models fail.

### 2014-12-03 writing/F'14 paper/quantifiers.tex
Intro update and misc paper cleanup
Tentative results in paper

We evaluate the models by randomly sampling 20\% of the data to
test on and training on the remainder. In this setting, each model being asked to learn a
complete reasoning system for the limited language and logic presented
in the training data. To succeed on this task, the model must be able
to recognize all of the lexical relations between the nouns, verbs,
and quantifiers and how they interact. For instance, it might see
\eqref{p1} and \eqref{p2} in training and, from that information,
determine \eqref{p3}.

@@

\begin{table}[tp]
  \centering \small
  \begin{tabular}{ l r@{ \ }r r@{ \ }r r@{ \ }r }
    \toprule
    ~&\multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{25d RNN}  & \multicolumn{2}{c}{25d RNTN}\\
    \midrule
    Test & 35.4 &(7.5) &	99.1&(99.2)& \textbf{99.5} & \textbf{(99.3)}\\
    Train & 35.4 &(7.5) &	100&		(100)&	100&	(100)\\	
    \bottomrule
  \end{tabular}
  \caption{Performance on the quantifier experiments, given as \% correct and macroaveraged F1.}
  \label{qresultstable}
\end{table} 
%
% do not allow a blank line --- adds too much space
%
\paragraph{Results} The results, reported in Figure \ref{qresultstable}, show that both models are able to learn to generalize the underlying logic well. We did not observe any consistent pattern in the few errors made by either model, and errors did not tend to be consistent across runs of the experiment.

>>>>

We evaluate the models by randomly sampling 20\% of the data to
test on and training on the remainder in five fold cross-validation.
In this setting, each model being asked to learn a
complete reasoning system for the limited language and logic presented
in the training data. To succeed on this task, the model must be able
to recognize all of the lexical relations between the nouns, verbs,
and quantifiers and how they interact. For instance, it might see
\eqref{p1} and \eqref{p2} in training and, from that information,
determine \eqref{p3}.

@@

\begin{table}[tp]
  \centering \small
  \begin{tabular}{ l r@{ \ }r r@{ \ }r r@{ \ }r }
    \toprule
    ~&\multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{25d RNN}  & \multicolumn{2}{c}{25d RNTN}\\
    \midrule
    Train & 35.4 &(7.5) &	100&		(100)&	100&	(100)\\	
    Test & 35.4 &(7.5) &	99.4&(99.4)& \textbf{99.5} & \textbf{(99.4)}\\
    \bottomrule
  \end{tabular}
  \caption{Performance on the quantifier experiments, given as \% correct and macroaveraged F1.}
  \label{qresultstable}
\end{table} 
%
% do not allow a blank line --- adds too much space
%
\paragraph{Results} The results, reported in Figure \ref{qresultstable}, show that both models are able to learn to generalize the underlying logic satisfactorily. We did not observe any consistent pattern in the few errors made by either model, and errors did not tend to be consistent across runs of the experiment, suggesting that there is no fundamental obstacle to learning a perfect model for this problem.

### 2014-12-03 writing/F'14 paper/recursion.tex
Tentative results in paper

\begin{figure*}[t]
  \centering
  \includegraphics[width=5.3in]{recursion\string_results\string_final.eps}
%  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
  \caption{\mynote{Model performance on propositional logic, by expression size. }
%    The dashed line indicates that only expressions of size four or less appeared in the training data. \textbf{Bottom:} Semantically distinct formulae should have different
%    representations. As measured by Euclidean distance, only the RNTN
%    achieves this for formulae containing more than a small number of
%    connectives (\ii{and} in this example). The RNN quickly collapses the representations,
%    failing to capture the meaning contrasts.
    }  
  \label{prop-results}
  % TODO: Regenerate figure from final model.
\end{figure*}

\paragraph{Results} 
We trained both the RNN and RNTN models on data with up to 4
connectives (60k pairs) and tested it on a separate set of examples of up to 12
connectives (21k pairs). We initialized the model parameters randomly,
including the vector representations of the six variables, and tested the model
in five fold cross-validation, discarding all but 20\% of the examples with 
more than four connectives. The results are shown in Figure~\ref{prop-results}. 
\mynote{In tuning, we found that the
RNN model was approximately optimal with 45-dimensional vector
representations, and the RNTN model was approximately optimal with 25
dimensions. We fixed the size of the feature vector for the classifier
at 75 dimensions. We found that the RNTN model was able to perform
almost perfectly on unseen small test examples, with accuracy above
98.7\% on formulae below length four (macro-F1 was above 95.5).
Starting at size four, performance gradually falls with increasing
size.  The RNN model did not perform well, reaching only 90.5\% (81.4
macro-F1) accuracy on the smallest test examples, and declining from
there to near-baseline performance on formulae of length 12. Training
accuracy was 99.6\% for the RNTN and 87.6\% for the RNN.}

The performance of the RNTN model on small unseen test examples
indicates that it learned a correct approximation of the underlying
logic. It appears that this approximation is accurate enough to yield
correct answers when the composition layer is only applied a small
number of times, but that the error in the approximation grows with
increasing expression size, resulting in gradually dropping performance. In
contrast, the RNN model did not appear to be able to accurately handle statements of any length, but it
appears that the approximation that the RNN did learn decayed similarly
with expression size.

\mynote{In an effort to better understand why the RNTN outperformed the plain
RNN, we evaluated both models on pairs of long formulae involving
binary connectives, to assess how well they distinguish representations for semantically distinct
formulae. We found that only the RNTN is able to learn
substantially different representations for pairs of differing formulae like $(a \pland (a \pland a))$
 and $(a \pland (a \pland b))$ when the difference between the two is placed under multiple operators.
The RNN separates the bare symbols by a euclidean distance of 3.5, but this falls to less than one once the two are placed under three operators, and quickly approaches zero as depth increases.
 The RNTN separates the bare symbols by 2.5, and this falls off much more gradually to 1.3 with twelve operators.}

 % TODO: If results are still bad tomorrow, write something about why this is not fatal to the program.

>>>>

\begin{figure*}[t]
  \centering
  \includegraphics[width=5.3in]{decayfig.eps}
%  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
  \caption{\mynote{Model performance on propositional logic, by expression size. }
%    The dashed line indicates that only expressions of size four or less appeared in the training data. \textbf{Bottom:} Semantically distinct formulae should have different
%    representations. As measured by Euclidean distance, only the RNTN
%    achieves this for formulae containing more than a small number of
%    connectives (\ii{and} in this example). The RNN quickly collapses the representations,
%    failing to capture the meaning contrasts.
    }  
  \label{prop-results}
  % TODO: Regenerate figure from final model.
\end{figure*}

\paragraph{Results} 
We trained both the RNN and RNTN models on data with up to 4
connectives (60k pairs) and tested it on a separate set of examples of up to 12
connectives (21k pairs). We initialized the model parameters randomly,
including the vector representations of the six variables, and tested the model
in five fold cross-validation, discarding all but 20\% of the examples with 
more than four connectives. The results are shown in Figure~\ref{prop-results}. 
In tuning, we found that the
RNN model was approximately optimal with 45-dimensional vector
representations, \mynote{and the RNTN model was approximately optimal with 25
dimensions. We found that the both models model were able to perform
well on unseen small test examples, with RNN accuracy above
97\% and RNTN accuracy above 98\% on formulae below length five.
Starting at size four, performance gradually falls with increasing
size. Training accuracy was 99\% for the RNN and 99.5\% for the RNTN.}

The performance of these model on small unseen test examples
indicates that it learned a correct approximation of the underlying
logic. It appears that this approximation is accurate enough to yield
correct answers when the composition layer is only applied a small
number of times, but that the error in the approximation grows with
increasing expression size, resulting in gradually dropping performance.

\mynote{In an effort to better understand how the models are decaying, 
we evaluated both models on pairs of long formulae involving
binary connectives, to assess how well they distinguish representations for semantically distinct
formulae. We found that only the RNTN is able to learn
substantially different representations for pairs of differing formulae like $(a \pland (a \pland a))$
 and $(a \pland (a \pland b))$ when the difference between the two is placed under multiple operators.
The RNN separates the bare symbols by a euclidean distance of 3.5, but this falls to less than one once the two are placed under three operators, and quickly approaches zero as depth increases.
 The RNTN separates the bare symbols by 2.5, and this falls off much more gradually to 1.3 with twelve operators.}

### 2014-12-03 writing/F'14 paper/sick.tex
Intro update and misc paper cleanup
Tentative results in paper

We also supplemented the SICK training data\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the SICK test set.} with \mynote{TODO} examples of entailment data from the Denotation Graph project \cite{hodoshimage}, a corpus of automatically labeled (and thus noisy) entailment examples over image captions, the same genre of text from which SICK was drawn. We trained a single model on data from both sources, but used a separate set of softmax parameters for classifying into the labels from each source. We parsed the data from both sources with the Stanford PCFG Parser v. 3.3.1 \cite{klein2003accurate}. We also found that we were able to train a working model much more quickly with an additional technique: we collapse subtrees that were identical across both sentences in a pair down to a single head word. Finally, in order to improve regularization on the noisier data, we used dropout \cite{hinton2012improving} at the input to the comparison layer (10\%) and at the output from the embedding transform layer (25\%).

% TODO: JMLR version of hinton paper.
% TODO: Add Karpathy to TACL paper

\begin{table}[tp]
  \centering \small
  \begin{tabular}{ l r@{ \ } r@{ \ } r@{ \ } }
    \toprule
    ~&\multicolumn{1}{c}{$\natind$ only} & \multicolumn{1}{c}{50d RNN}  & \multicolumn{1}{c}{50d RNTN}\\
    \midrule
    Test & 56.7 &	...& \textbf{75.5} \\
    Train &- &- &-  \\
    \bottomrule
  \end{tabular}
  \caption{\mynote{Performance on SICK. Training figures include performance on the Denotation Graph data.}}
  \label{sresultstable}
\end{table}

Despite the small amount of high quality training data available and the lack of resources for learning lexical relationships (especially exclusion, as in \ii{cat}--\ii{dog}), it is possible to train our model to perform competitively on textual entailment. Our performance did not reach the state of the art (84.6\%), but exceeded the median submission to the competition (77.06\%) in a field including many models which used sophisticated hand-engineered features and lexical resources specific to the version of the entailment task at hand.

>>>>

We also supplemented the SICK training data\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the SICK test set. We also report training accuracy on a small sample from each data source.} with 600k examples of entailment data from the Denotation Graph project (DG, \citealt{hodoshimage}), a corpus of automatically labeled (and thus noisy) entailment examples over image captions, the same genre of text from which SICK was drawn. We trained a single model on data from both sources, but used a separate set of softmax parameters for classifying into the labels from each source. We parsed the data from both sources with the Stanford PCFG Parser v. 3.3.1 \cite{klein2003accurate}. We also found that we were able to train a working model much more quickly with an additional technique: we collapse subtrees that were identical across both sentences in a pair down to a single head word. Finally, in order to improve regularization on the noisier data, we used dropout \cite{srivastava2014dropout} at the input to the comparison layer (10\%) and at the output from the embedding transform layer (25\%).

\begin{table}[tp]
  \centering \small
  \begin{tabular}{ l r@{ \ } r@{ \ } r@{ \ } }
    \toprule
    ~&\multicolumn{1}{c}{\ii{neut.} only} & \multicolumn{1}{c}{30d RNN}  & \multicolumn{1}{c}{50d RNTN}\\
    \midrule
    SICK Train &  56.7 & \textbf{95.6} &  95.0  \\
    DG Train &  50.0 & \textbf{68.0} & 66.0  \\
    Test & 56.7 & 75.3 & \textbf{75.7}  \\
    \bottomrule
  \end{tabular}
  \caption{\mynote{Performance on SICK.}}
  \label{sresultstable}
\end{table}

The results are shown in Table \ref{sresultstable}. Despite the small amount of high quality training data available and the lack of resources for learning lexical relationships (especially exclusion, as in \ii{cat}--\ii{dog}), it is possible to train our model to perform competitively on textual entailment. Our performance did not reach the state of the art (84.6\%), but exceeded the performance of eight out of 18 submitted systems including several which used sophisticated hand-engineered features and lexical resources specific to the version of the entailment task at hand.
